---
title: "Exploring trends and variation in biologic therapies for severe asthma"
# description: "Secondary Care Medicines Data"
date: "`r lubridate::today()`"
repository_url: https://github.com/ebmdatalab/open-nhs-hospital-medicines-pilot
output:
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float: true
navbar:
  search: true
bibliography: "references.bib"
link-citations: true
editor_options: 
  chunk_output_type: console
---

```{r rmd-setup, include=FALSE}
# R Markdown settings
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```


```{r setup, include=FALSE}
# Load utils for this project
source(here::here("scripts/connect_bigquery.R"))
source(here::here("scripts/sql_query_scmd.R"))
source(here::here("scripts/sql_query_dmd.R"))
# DBI::dbListTables(conn_ebm_scmd)

# Load R packages
library(tidyverse)
library(bigrquery)
library(DBI)
library(here)
library(reactable)
library(htmltools)
library(plotly)
library(dplyr)
library(dbplyr)
library(sparkline)
library(tsibble)
library(lubridate)
library(tippy)
```

# Background

Patients with severe asthma can be treated with biologic medicinces. 
NICE has approved four biologics for use in England for people with severe asthma; benralizumab, mepolizumab and most recently reslizumab for treating severe eosinophilic asthma and omalizumab for treating severe persistent confirmed allergic IgE‑mediated asthma. 
Our team delivers [OpenPrescribing.net](https://openprescribing.net/), a publicly funded and openly accessible explorer for NHS primary care prescribing supported by prescribing data openly published by the NHS Business Services Authority. 
We were concerned that the NHS did not share hospital medicines data in a similar manner to primary care and that it should be shared [see @Goldacre2020]. 
Asthma UK have [similarly advocated](https://www.asthma.org.uk/support-us/campaigns/publications/living-in-limbo/) that hospital medicines data should be shared to enable transparent assessment of access to new treatments and accountability for public spending.

In September 2020 NHS BSA published hospital medicines data for the first time. 
We have prepared the following notebook for investigating the use of asthma biologic medicines in English hospitals. 
If you have any feedback or insight The DataLab team can be contacted at ebmdatalab@phc.ox.ac.uk.

# Methods

## Medicines codelist

The list `snomed_asthma_mabs_list` contains the SNOMED codes for the medicines we are investigating.

```{r}
# Define vector with selected SNOMED codes for Asthma Medicines
# Consider getting these from opencodelists at some point maybe
snomed_asthma_mabs_list <- c("18671311000001102",
                             "18671411000001109",
                             "37564311000001105",
                             "37564411000001103",
                             "31210311000001108",
                             "33999611000001101",
                             "34812511000001102",
                             "37854611000001108",
                             "35298511000001102")
```

## Data

Before we can analyse trends and variation in the use of asthma biologics we need to prepare our dataset. 
This analysis uses three differen sources of data.


### NHS Trusts

- The NHS Digital "GP mapping file" which provides STP and region names mapped to STP and region ODS codes, published [here](https://digital.nhs.uk/data-and-information/publications/statistical/patients-registered-at-a-gp-practice/september-2020).

- The NHS Digital "Etr" file that maps trust organisation codes to trust names, STP ODS codes and region ODS codes, published [here](https://digital.nhs.uk/services/organisation-data-service/data-downloads/other-nhs-organisations).


```{r, cache=TRUE}
# Load ETR data
df_etr <- readr::read_csv(here::here("data/etr_tidy.csv")) 

# Load stp to regions data
stp_to_region_map <- read_csv(here::here("data/gp-reg-pat-prac-map.csv")) %>%
  group_by(STP_CODE, STP_NAME) %>%
  summarise(COMM_REGION_NAME = first(COMM_REGION_NAME),
            COMM_REGION_CODE = first(COMM_REGION_CODE)) %>%
  janitor::clean_names()

```

```{r tab-etr, cache=TRUE}
# Sustainability and Transformation Partnerships (STPs) 
df_etr %>% 
  left_join(stp_to_region_map) %>% 
  select(ods_name, stp_name, comm_region_name) %>% 
  mutate(stp_name = fct_explicit_na(stp_name),
         comm_region_name = fct_explicit_na(comm_region_name)) %>%
  reactable::reactable(filterable = TRUE,
                       columns = list(ods_name = reactable::colDef(name = "Name", 
                                                                   minWidth = 200),
                                      stp_name = reactable::colDef(name = "STP", 
                                                                   minWidth = 150),
                                      comm_region_name = reactable::colDef(name = "Region", 
                                                                           minWidth = 70)
                       ),
                       style = list(fontSize = "12px"),
                       highlight = TRUE)
```

### SCMD

- The Secondary Care Medicines Dataset (SCMD) published by the NHS BSA, see [here](https://opendata.nhsbsa.net/dataset/d5d7cd63-5373-4a0e-ae22-7f1a198e305d).

```{r, cache=TRUE}
# Secondary Care Medicines Data
# Connect to database, filter, and collect data.
# Get SCMD dataset
# Check negative quantities, seems to be a small (0.7%) problem in the data
# dplyr::tbl(conn_ebm_scmd, "scmd") %>% 
#   filter(between(year_month, "2019-01-01", "2019-12-31")) %>% 
#   # arrange(desc(year_month))
#   filter(total_quanity_in_vmp_unit > 0) %>% 
#   count()
#   group_by(vmp_snomed_code) %>% 
#   summarise(n = n(),
#             sum = sum(total_quanity_in_vmp_unit < 0))

db_scmd <- dplyr::tbl(conn_ebm_scmd, sql_query_scmd)

# Create dataframe for table
db_scmd_asthma <- db_scmd %>% 
  dplyr::filter(vmp_snomed_code %in% snomed_asthma_mabs_list)

df_scmd_asthma <- dplyr::collect(db_scmd_asthma)

# df_scmd_asthma %>% skimr::skim()

# Tidy tidy tidy data
df_scmd_asthma_names <- df_scmd_asthma %>% 
  dplyr::left_join(dplyr::select(df_etr, ods_code, ods_name, stp_code), by = "ods_code") %>% 
  # some data cleaning as scmd uses some ods codes that are not up to date
  mutate(stp_code = as.character(stp_code),
         stp_code = case_when(
           ods_code == "RQ6" ~ "QYG",
           ods_code %in% c("RNL", "RE9", "RLN") ~ "QHM",
           ods_code %in% c("RM2", "RW3") ~ "QOP",
           ods_code == "RGQ" ~ "QJG",
           ods_code == "RJF" ~ "QJ2",
           ods_code == "RR1" ~ "QHL",
           TRUE ~ stp_code
         ))
```


```{r}
# Fill explicit missing and create dataset for sparkline in table
df_tab_sparkline <- df_scmd_asthma_names %>% 
  select(-c(vmp_product_name, ods_name, stp_code, stp_code, ods_name)) %>% 
  arrange(ods_code, vmp_snomed_code, year_month) %>% 
  as_tsibble(key = c(ods_code, vmp_snomed_code), index = year_month) %>% 
  fill_gaps(total_quantity = 0, .full = TRUE) %>% 
  tidyr::fill(.direction = "down") %>% 
  as_tibble() %>% 
  mutate(year_month = floor_date(year_month, unit = "month")) %>% 
  group_by(year_month, ods_code, vmp_snomed_code) %>% 
  arrange(ods_code, vmp_snomed_code, year_month) %>% 
  mutate(total_quantity = sum(total_quantity)) %>%
  arrange(ods_code, vmp_snomed_code, year_month) %>% 
  distinct() %>% 
  group_by(ods_code, vmp_snomed_code) %>%
  dplyr::summarise(count_sparkline = list(total_quantity)) %>% 
  group_by(ods_code, vmp_snomed_code) %>% 
  dplyr::mutate(total_quantity = sum(unlist(count_sparkline)))

# Create lookup datasets for joining
# SNOMED
vmp_snomed_names_lookup <- df_scmd_asthma_names %>% 
  select(vmp_snomed_code, vmp_product_name) %>% 
  dplyr::distinct()

# Trust
trust_names_lookup <- df_scmd_asthma_names %>% 
  select(ods_code, ods_name, stp_code) %>% 
  dplyr::distinct()

# Join data
df_tab_sparkline <- df_tab_sparkline %>%
  ungroup() %>% 
  arrange(ods_code) %>% 
  left_join(trust_names_lookup, by = c("ods_code")) %>% 
  left_join(vmp_snomed_names_lookup, by = c("vmp_snomed_code")) %>% 
  mutate(count_box = count_sparkline)

# See the ?tippy documentation to learn how to customize tooltips
with_tooltip <- function(value, tooltip, ...) {
  div(style = "text-decoration: underline; text-decoration-style: dotted; cursor: help",
      tippy(value, tooltip, ...))
}
# Create table
df_tab_sparkline %>% 
  select(ods_name, vmp_product_name,vmp_snomed_code, count_sparkline, count_box, total_quantity, 
         -ods_code, -stp_code) %>% 
  reactable(filterable = TRUE,
            defaultSorted = c("ods_name", "total_quantity"),
            groupBy = c("ods_name"),
            columns = list(
              ods_name = reactable::colDef(name = "Trust", 
                                           # header = with_tooltip("Trust", "ods_name"),
                                           minWidth = 200),
              count_sparkline = colDef(name = "Trend",
                                       header = with_tooltip("Trend", "Note that the y axis cannot be compared across different entries."),
                                       minWidth = 50,
                                       cell = function(value, index) {
                                         sparkline(df_tab_sparkline$count_sparkline[[index]])
                                       }),
              # count_box = colDef(name = "Boxplot",
              #                    header = with_tooltip("Boxplot", "Boxplot of monthly quantity for entries with more than 10 data points."),
              #                    minWidth = 40,
              #                    cell = function(value, index) {
              #                      if (sum(df_tab_sparkline$count_box[[index]] > 0) > 10) {
              #                        
              #                        sparkline(df_tab_sparkline$count_box[[index]],
              #                                  type = "box")
              #                      }
              #                    }),
              count_box = reactable::colDef(show = FALSE),
              total_quantity = reactable::colDef(name = "Quantity",
                                                 # header = with_tooltip("Total quantity", "total_quantity"),
                                                 minWidth = 50,
                                                 aggregate = "sum",
                                                 format = reactable::colFormat(digits = 0)),
              vmp_product_name = reactable::colDef(name = "Product", 
                                                   # header = with_tooltip("Product", "total_quantity"),

                                                   minWidth = 150,
                                                   cell = function(value, index) {
                                                     vmp_snomed_code <- paste0("SNOMED: ", df_tab_sparkline$vmp_snomed_code[index])
                                                     vmp_snomed_code <- if (!is.na(vmp_snomed_code)) vmp_snomed_code else "Unknown"
                                                     div(
                                                       div(style = list(fontWeight = 600), value),
                                                       div(style = list(fontSize = 10), vmp_snomed_code))
                                                   }
              ),
              vmp_snomed_code = reactable::colDef(show = FALSE)
            ),
            style = list(fontSize = "12px"),
            highlight = TRUE
)

```

### dm+d

Information from the dm+d (Dictionary of Medicines and Devices) on the strength and vmp quantity of each asthma biologic at VMP level, using data hosted on the DataLab BigQuery server.

```{r, cache=TRUE}
db_dmd_info <- dplyr::tbl(conn_ebm_scmd, sql_query_dmd_info)

df_dmd_info <- db_dmd_info %>% 
  filter(vmp_snomed_code %in% snomed_asthma_mabs_list) %>% 
  collect()

```

We use information on the daily defined dose (DDD) of each asthma biologic, so that the volume of each VMP can be compared directly once converted to DDDs. The WHO publish DDDs online:

- [Omalizumab DDD](https://www.whocc.no/atc_ddd_index/?code=R03DX05) = 16mg 
- [Mepolizumab DDD](https://www.whocc.no/atc_ddd_index/?code=R03DX09) = 3.6mg
- [Reslizumab DDD](https://www.whocc.no/atc_ddd_index/?code=R03DX08) = 7.1mg
- [Benralizumab DDD](https://www.whocc.no/atc_ddd_index/?code=R03DX10) = 0.54mg 

```{r}
# Define tibble with mg_per_ddd for join later
# Is there a way this step could be automated?
lookup_mg_per_ddd <- tibble::tribble(~vtmnm, ~mg_per_ddd,
                                     "Omalizumab", 16, 
                                     "Mepolizumab", 3.6, 
                                     "Reslizumab", 7.1, 
                                     "Benralizumab", 0.54)

df_scmd_asthma_mg <- df_scmd_asthma_names %>% 
  left_join(df_dmd_info, by = c("vmp_snomed_code", "vmp_product_name")) %>% 
  left_join(lookup_mg_per_ddd, by = "vtmnm") 
```


```{r}
df_scmd_asthma_mg %>% 
  select(vmp_snomed_code, vtmnm, form_descr, udfs, udfs_descr,
         strnt_nmrtr_val, strnt_nmrtr_uom, strnt_dnmtr_val,strnt_dnmtr_descr, mg_per_ddd) %>% 
  distinct() %>% 
    reactable(filterable = TRUE,
              columns = list(
                vmp_snomed_code = reactable::colDef(name = "SNOMED", 
                                                    minWidth = 100),
                vtmnm = reactable::colDef(name = "Name", 
                                          minWidth = 100),
                form_descr = reactable::colDef(name = "Form", 
                                               minWidth = 80),
                # udfs is the VMP unit dose form strength
                udfs = reactable::colDef(name = "Value", 
                                         minWidth = 40),
                udfs_descr = reactable::colDef(name = "Unit", 
                                               minWidth = 40),
                # strnt_nmrtr
                strnt_nmrtr_val = reactable::colDef(name = "Numerator", 
                                                    minWidth = 60,
                                                    format = colFormat(suffix = " mg")),
                strnt_nmrtr_uom = reactable::colDef(show = FALSE),
                # strnt_dnmtr
                strnt_dnmtr_val = reactable::colDef(name = "Denominator", 
                                                    minWidth = 60,
                                                    format = colFormat(suffix = " ml")),
                strnt_dnmtr_descr = reactable::colDef(show = FALSE),
                mg_per_ddd = reactable::colDef(name = "mg/ddd", 
                                               minWidth = 50)),
              columnGroups = list(
                colGroup(name = "UDFS", columns = c("udfs", "udfs_descr")),
                colGroup(name = "Strength", columns = c("strnt_nmrtr_val", "strnt_dnmtr_val"))
              ),
              style = list(fontSize = "12px"),
              highlight = TRUE
    )


```

## Convert volume

The final data cleaning step is to convert the volume from VMP quantity (as provided in the SCMD dataset) to volume in DDDs.

- SCMD volumes data is provided in vmp quantity - this means different things for different products. e.g. 100mg powder = 1 vmp quantity but 100mg/20ml solution for injection = 20 vmp quantity, even though both VMPS have the same strength of ingredient.
- To translate volume in VMP quantity to volume in DDDs we need to go through a few steps:
  - First, translate volume in VMP quantity to volumes in singles of the product (i.e. number of vials)
  - Then translate volume in singles of product to volume in strength of ingredient (i.e. number of mgs of active ingredient)
  - Finally translate volume in strength of ingredient to volume in DDDs, using the DDD information published by the WHO

```{r}

df_scmd_asthma_ddd <- df_scmd_asthma_mg %>% 
  mutate(volume_singles = total_quantity / udfs,
         volume_mg_strength = volume_singles * if_else(is.na(strnt_dnmtr_val), 
                                                       true = strnt_nmrtr_val, 
                                                       false = strnt_nmrtr_val * 
                                                               (udfs / strnt_dnmtr_val)),
         volume_ddd = volume_mg_strength / mg_per_ddd)

```





# Results

## Trends over time

### National prescribing

```{r, fig.width=7.2, fig.height=3, cache=TRUE, fig.cap="Figure. National prescribing of asthma biologics over time."}

temp_ggplot <- df_scmd_asthma_ddd %>%
  group_by(year_month, vtmnm) %>%
  summarise(volume_ddd = sum(volume_ddd)) %>% 
  ggplot(aes(x = year_month, 
             y = volume_ddd, 
             colour = vtmnm, 
             group = vtmnm)) +
  geom_line(size = 1, alpha = 0.5) +
  geom_point(aes(text = paste0("<b>Month:</b> ", 
                               lubridate::month(year_month, label = TRUE), " ",
                               lubridate::year(year_month), "<br>",
                               "<b>Volume:</b> ", round(volume_ddd, 0), "<br>",
                               "<b>Medication:</b> ", vtmnm)), size = 2) +
  scale_x_date(date_breaks = "4 month", date_labels =  "%b %y") +  scale_colour_viridis_d() +
  labs(x = NULL, y = "Defined Daily Dose",
       colour = NULL) +
  geom_vline(xintercept = as.numeric(as.Date("2020-03-31")), 
             color = "orange", 
             linetype = 2,
             lwd = .5, 
             alpha = .5) +
  theme(text = element_text(size = 12))

# temp_ggplot

plotly::ggplotly(temp_ggplot,
                 tooltip = "text") %>%
  plotly::config(displayModeBar = FALSE)
```


### Regional prescribing

```{r}
df_scmd_asthma_ddd_map <- df_scmd_asthma_ddd %>% 
  left_join(stp_to_region_map, by = "stp_code")
```



```{r, fig.width=8, fig.height=8, fig.cap="Figure. Regional prescribing of asthma biologics over time."}

temp_ggplot <- df_scmd_asthma_ddd_map %>%
  group_by(comm_region_name, year_month, vtmnm) %>%
  summarise(volume_ddd = sum(volume_ddd)) %>% 
  mutate(comm_region_name = fct_explicit_na(comm_region_name)) %>% 
  ggplot(aes(x = year_month, y = volume_ddd, 
             colour = comm_region_name, group = comm_region_name)) +
  geom_vline(xintercept = as.numeric(as.Date("2020-03-31")), 
             color = "orange", 
             linetype = 2,
             lwd = .5, 
             alpha = .5) +
  geom_line(size = 1, alpha = 0.5) +
  geom_point(aes(text = paste0("<b>Month:</b> ", 
                               lubridate::month(year_month, label = TRUE), " ",
                               lubridate::year(year_month), "<br>",
                               "<b>Region:</b> ", comm_region_name, "<br>",
                               "<b>Volume:</b> ", round(volume_ddd, 0), "<br>",
                               "<b>Medication:</b> ", vtmnm)), 
             size = 2) +
  scale_x_date(date_breaks = "4 month", date_labels =  "%b %y") +
  scale_colour_viridis_d(end = 1) +
  labs(x = NULL, 
       y = "Defined Daily Dose",
       colour = NULL) +
  facet_wrap(~vtmnm, ncol = 1) +
  theme(text = element_text(size = 12)) 
  
# temp_ggplot

plotly::ggplotly(temp_ggplot,
                 tooltip = "text") %>%
  plotly::config(displayModeBar = FALSE)

```

## Total volume per year

### Regional prescribing

```{r fig.width=8, fig.height=4, fig.cap="Figure. Total regional prescribing of asthma biologics."}
temp_ggplot <- df_scmd_asthma_ddd_map %>% 
  select(year_month, ods_code, vtmnm, volume_ddd, comm_region_name) %>% 
  mutate(comm_region_name = fct_explicit_na(comm_region_name)) %>% 
  group_by(comm_region_name, vtmnm) %>%
  summarise(volume_ddd = sum(volume_ddd)) %>%
  group_by(comm_region_name) %>%
  mutate(prop_use = volume_ddd / sum(volume_ddd),
         pos = cumsum(volume_ddd) - volume_ddd/2,
         total = sum(volume_ddd)) %>%
  ungroup() %>% 
  mutate(comm_region_name = fct_reorder(comm_region_name, total)) %>% 
  ggplot(aes(comm_region_name)) +
  geom_bar(aes(y = volume_ddd,
               fill = vtmnm,
               text = paste0("<b>Region:</b> ", comm_region_name, "<br>",
                             "<b>Total volume in ddd:</b> ", round(total, 0), "<br>",
                             # "<b>Medication:</b> ", vtmnm, "<br>",
                             "<b>", vtmnm , " volume in ddd (%):</b> ", round(volume_ddd, 0), " (", scales::percent(prop_use, accuracy = 0.1), ")"
                             )
               ), 
           stat='identity',
           # position = position_dodge()
           ) +
  scale_fill_viridis_d() +
  labs(subtitle = paste0("From: ", min(df_scmd_asthma_ddd_map$year_month), " to ", max(df_scmd_asthma_ddd_map$year_month)),
       x = NULL,
       y = "Defined Daily Dose",
       fill = NULL) +
  scale_y_continuous(labels = scales::comma) +
  theme(text = element_text(size = 12)) +
  coord_flip()

# temp_ggplot
plotly::ggplotly(temp_ggplot,
                 tooltip = "text") %>%
  plotly::config(displayModeBar = FALSE)
```

### STPs

```{r, fig.width=8, fig.height=6, fig.cap="Figure. Total prescribing of asthma biologics for 20 STPs with the largest volume across all selected medications."}
df_scmd_asthma_ddd_map_temp <- df_scmd_asthma_ddd_map %>% 
  filter(year_month >= as.Date("2019-08-01") & year_month <= as.Date("2020-07-01")) %>%
  select(year_month, ods_code, vtmnm, volume_ddd, stp_name) %>% 
  mutate(stp_name = fct_explicit_na(stp_name)) %>% 
  group_by(stp_name, vtmnm) %>%
  summarise(volume_ddd = sum(volume_ddd)) %>%
  group_by(stp_name) %>%
  mutate(prop_use = volume_ddd / sum(volume_ddd),
         pos = cumsum(volume_ddd) - volume_ddd/2,
         total = sum(volume_ddd)) %>%
  ungroup() %>% 
  mutate(rank = dense_rank(-total),
         stp_name = fct_reorder(stp_name, -rank))

temp_ggplot <- df_scmd_asthma_ddd_map_temp %>% 
  filter(rank <= 20) %>% 
  ggplot(aes(stp_name)) +
  geom_bar(aes(y = volume_ddd,
               fill = vtmnm,
               text = paste0("<b>STP:</b> ", stp_name, "<br>",
                             "<b>Total volume in ddd:</b> ", round(total, 0), "<br>",
                             # "<b>Medication:</b> ", vtmnm, "<br>",
                             "<b>", vtmnm , " volume in ddd:</b> ", round(volume_ddd, 0), " (", scales::percent(prop_use, accuracy = 0.1), ")"
               )
               ),
               # text = paste0("<b>STP:</b> ", stp_name, "<br>",
               #               "<b>Volume:</b> ", round(total, 0), "<br>",
               #               "<b>Proportion:</b> ", scales::percent(prop_use, accuracy = 0.1), "<br>",
               #               "<b>Medication:</b> ", vtmnm)), 
           stat='identity',
           # position = position_dodge()
           ) +
  scale_fill_viridis_d() +
  labs(subtitle = paste0("From: ", min(df_scmd_asthma_ddd_map$year_month), " to ", max(df_scmd_asthma_ddd_map$year_month)),
       x = NULL,
       y = "Defined Daily Dose",
       fill = NULL) +
  scale_y_continuous(labels = scales::comma) +
  theme(text = element_text(size = 12),
        legend.position="bottom") +
  coord_flip()

# temp_ggplot
plotly::ggplotly(temp_ggplot,
                 tooltip = "text") %>%
  plotly::config(displayModeBar = FALSE) %>% 
layout(legend = list(orientation = "h", x = -0.5, y =-.15))
```



```{r}
green_pal <- function(x) rgb(colorRamp(c("#effbf7", "#31cf96"))(x), maxColorValue = 255)

df_scmd_asthma_ddd_map_temp_temp <- df_scmd_asthma_ddd_map_temp %>% 
  mutate(prop_use = round(prop_use, 3)) %>% 
  pivot_wider(id_cols = c(stp_name, total, rank),
              names_from = vtmnm, 
              values_from = c(volume_ddd, prop_use), 
              values_fill = 0) %>% 
  select(stp_name, starts_with("prop"), total)

data_tab <- df_scmd_asthma_ddd_map_temp_temp %>% 
  mutate(prop_scale = seq(from = 0, to = 1, length = nrow(df_scmd_asthma_ddd_map_temp_temp)))

reactable(data_tab, 
          filterable = TRUE,
          # searchable = TRUE,
          columns = list(
            stp_name = colDef(name = "STP", 
                              minWidth = 300),
            prop_use_Benralizumab = colDef(name = "Benralizumab", 
                                           format = colFormat(percent = TRUE, digits = 1),
                                           minWidth = 100,
                                           style = function(value) {
                                             normalized <- (value - min(data_tab$prop_scale)) / (max(data_tab$prop_scale) - min(data_tab$prop_scale))
                                             color <- green_pal(normalized)
                                             list(background = color)
                                           }),
            prop_use_Mepolizumab = colDef(name = "Mepolizumab", 
                                          format = colFormat(percent = TRUE, digits = 1),
                                          minWidth = 100,
                                           style = function(value) {
                                             normalized <- (value - min(data_tab$prop_scale)) / (max(data_tab$prop_scale) - min(data_tab$prop_scale))
                                             color <- green_pal(normalized)
                                             list(background = color)
                                           }),
            prop_use_Omalizumab = colDef(name = "Omalizumab", 
                                         format = colFormat(percent = TRUE, digits = 1),
                                         minWidth = 100,
                                           style = function(value) {
                                             normalized <- (value - min(data_tab$prop_scale)) / (max(data_tab$prop_scale) - min(data_tab$prop_scale))
                                             color <- green_pal(normalized)
                                             list(background = color)
                                           }),
            prop_use_Reslizumab = colDef(name = "Reslizumab", 
                                         format = colFormat(percent = TRUE, digits = 1),
                                         minWidth = 100,
                                           style = function(value) {
                                             normalized <- (value - min(data_tab$prop_scale)) / (max(data_tab$prop_scale) - min(data_tab$prop_scale))
                                             color <- green_pal(normalized)
                                             list(background = color)
                                           }),
            total = colDef(name = "Total (ddd)", 
                           minWidth = 65,
                           format = colFormat(digits = 0)),
            prop_scale = reactable::colDef(show = FALSE)
            ),
          style = list(fontSize = "12px"),
          highlight = TRUE)

```


# References
