---
title: "Asthma Monoclonal Antibodies"
author: "Anna Rowan"
date: "08/10/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{css, echo=FALSE}
pre {
  max-height: 300px;
  overflow-y: auto;
}

pre[class] {
  max-height: 200px;
}
```


## Introduction

Patients with severe asthma can be treated with biologic medicinces. NICE has approved four biologics for use in England for people with severe asthma; benralizumab, mepolizumab and most recently reslizumab for treating severe eosinophilic asthma and omalizumab for treating severe persistent confirmed allergic IgE‑mediated asthma.

In September 2020 NHS BSA published hospital medicines data for the first time. This data allows us to analyse the use of asthma monoclonal antibodies (asthma mAbs) in hospitals in England.

## Data preperation

Before we can analyse trends and variation in the use of asthma mAbs we need to prepare our dataset. This analysis uses:

* The secondary care medicines dataset, published by the NHS BSA [here.](https://opendata.nhsbsa.net/dataset/d5d7cd63-5373-4a0e-ae22-7f1a198e305d)

* The NHS Digital "Etr" file that maps trust organisation codes to trust names, STP ODS codes and region ODS codes, published [here.](https://digital.nhs.uk/services/organisation-data-service/data-downloads/other-nhs-organisations)

* The NHS Digital "GP mapping file" which provides STP and region names mapped to STP and region ODS codes, published [here.](https://digital.nhs.uk/data-and-information/publications/statistical/patients-registered-at-a-gp-practice/september-2020)

* Information from the dm+d on the strength and vmp quantity of each asthma mAb at VMP level, using data hosted on the DataLab BigQuery server.

We also use information on the daily defined dose (DDD) of each asthma mAb, so that the volume of each VMP can be compared directly once converted to DDDs. The WHO publish DDDs online:

* [Omalizumab DDD](https://www.whocc.no/atc_ddd_index/?code=R03DX05) = 16mg 

* [Mepolizumab DDD](https://www.whocc.no/atc_ddd_index/?code=R03DX09) = 3.6mg

* [Reslizumab DDD](https://www.whocc.no/atc_ddd_index/?code=R03DX08) = 7.1mg

* [Benralizumab DDD](https://www.whocc.no/atc_ddd_index/?code=R03DX10) = 0.54mg 

### Reading in the secondary care medicines dataset from DataLab BigQuery

```{r, message = FALSE, warning = FALSE}
# reading in hospital issues data for asthma mabs
library(tidyverse)
library(bigrquery)
library(DBI)
library(knitr)

options(scipen = 999)

## set up connection to BigQuery
con <- dbConnect(
  bigrquery::bigquery(),
  project = "ebmdatalab",
  dataset = "scmd")

scmd_extract_asthma_mabs <- "select 
year_month,
ods_code,
vmp_snomed_code,
vmp_product_name,
sum(total_quanity_in_vmp_unit) as total_quantity
from ebmdatalab.scmd.scmd
where 
vmp_snomed_code = \"18671311000001102\" or 
vmp_snomed_code = \"18671411000001109\" or
vmp_snomed_code = \"37564311000001105\" or
vmp_snomed_code = \"37564411000001103\" or
vmp_snomed_code = \"31210311000001108\" or
vmp_snomed_code = \"33999611000001101\" or
vmp_snomed_code = \"34812511000001102\" or
vmp_snomed_code = \"37854611000001108\" or
vmp_snomed_code = \"35298511000001102\"
group by
year_month,
ods_code,
vmp_snomed_code,
vmp_product_name
order by
year_month"

scmd_asthma_mabs <- dbGetQuery(con, scmd_extract_asthma_mabs)

kable(scmd_asthma_mabs[1:10, ], caption = "Preview of scmd_asthma_mabs data table")

```

### Reading in the dm+d information for asthma mAbs from DataLab BigQuery

```{r, message = FALSE, warning = FALSE}
# reading in dm+d data for asthma mabs

## set up connection to BigQuery
con <- dbConnect(
  bigrquery::bigquery(),
  project = "ebmdatalab",
  dataset = "scmd")

dmd_extract_asthma_mabs <- "select 
  cast(a.id as STRING) as vmpid,
  a.nm as vmpnm,
  a.vtm as vtmid,
  j.nm as vtmnm,
  b.form as form_cd,
  c.descr as formdescr,
  a.df_ind as df_ind_cd,
  d.descr as df_descr,
  a.udfs,
  e.descr as udfs_descr,
  f.descr as unit_dose_descr,
  g.strnt_nmrtr_val,
  h.descr as strnt_nmrtr_uom,
  g.strnt_dnmtr_val,
  i.descr as strnt_dnmtr_descr,
  a.bnf_code,
  k.presentation as bnf_presentation
  
from ebmdatalab.dmd.vmp as a

left join ebmdatalab.dmd.dform as b
on a.id = b.vmp

left join ebmdatalab.dmd.form as c
on b.form = c.cd

left join ebmdatalab.dmd.dfindicator as d
on a.df_ind = d.cd

left join ebmdatalab.dmd.unitofmeasure as e
on a.udfs_uom = e.cd

left join ebmdatalab.dmd.unitofmeasure as f
on a.unit_dose_uom = f.cd

left join ebmdatalab.dmd.vpi as g
on a.id = g.vmp

left join ebmdatalab.dmd.unitofmeasure as h
on g.strnt_nmrtr_uom = h.cd

left join ebmdatalab.dmd.unitofmeasure as i
on g.strnt_dnmtr_uom = i.cd

left join ebmdatalab.dmd.vtm as j
on a.vtm = j.id

left join ebmdatalab.hscic.bnf as k
on a.bnf_code = k.presentation_code

where
  a.id = 18671311000001102 or 
  a.id = 18671411000001109 or
  a.id = 37564311000001105 or
  a.id = 37564411000001103 or
  a.id = 31210311000001108 or
  a.id = 33999611000001101 or
  a.id = 34812511000001102 or
  a.id = 37854611000001108 or
  a.id = 35298511000001102"

dmd_asthma_mabs <- dbGetQuery(con, dmd_extract_asthma_mabs) %>% 
  dplyr::rename(vmp_product_name = vmpnm) %>%
  select(-c("vmpid"))

kable(dmd_asthma_mabs[1:5, ], caption = "Preview of dmd_asthma_mabs data table")

```

### Preview of hospital trust code to NHS Digital ODS information

```{r, message = FALSE, warning = FALSE}
# reading in ods data on hospitals - no header information so need to add this in
hospital_to_stp_map <- read.csv(file = here::here("data", "etr.csv"), header = FALSE) %>%
  select(1:4) 
colnames(hospital_to_stp_map) <- c("ods_code", "trust_name", "region_code", "stp_code")

kable(hospital_to_stp_map[1:5, ], caption = "Preview of hospital_to_stp_map data table")
```

```{r, message = FALSE, warning = FALSE}
# reading in map from ccg codes to stp and region codes because this gives a map from stp_code to stp_name
stp_to_region_map <- read.csv(here::here("data","gp-reg-pat-prac-map.csv")) %>%
  group_by(STP_CODE, STP_NAME) %>%
  summarise(COMM_REGION_NAME = first(COMM_REGION_NAME),
            COMM_REGION_CODE = first(COMM_REGION_CODE)) %>%
  dplyr::rename(stp_code = STP_CODE)

kable(stp_to_region_map[1:5, ], caption = "Preview of stp_to_region_map data table")
```

### Merging together and cleaning data sources

Now the four data tables will be merged together to create the dataset used for our analysis. To be able to match trust codes to trust names, STP ODS codes and region ODS codes there first needs to be some data cleaning as the SCMD uses some old trust ODS codes and doesn't take account of some trust mergers over the last few years.

```{r, message = FALSE, warning = FALSE}
# Merging data tables together
asthma_mabs <- left_join(scmd_asthma_mabs, dmd_asthma_mabs, by = "vmp_product_name") %>%
  left_join(., hospital_to_stp_map, by = "ods_code") %>%
  # some data cleaning as scmd uses some ods codes that are not up to date
  mutate(stp_code = as.character(stp_code),
         stp_code = case_when(
           ods_code == "RQ6" ~ "QYG",
           ods_code %in% c("RNL", "RE9", "RLN") ~ "QHM",
           ods_code %in% c("RM2", "RW3") ~ "QOP",
           ods_code == "RGQ" ~ "QJG",
           ods_code == "RJF" ~ "QJ2",
           ods_code == "RR1" ~ "QHL",
           TRUE ~ stp_code
         )
  ) %>%
  left_join(., stp_to_region_map, by = "stp_code") %>%
  select(-c("COMM_REGION_CODE"))

kable(asthma_mabs[1:5, ], caption = "Preview of asthma_mabs data table")
```

### Converting volume in VMP quantity to volume in DDDs

The final data cleaning step is to convert the volume from VMP quantity (as provided in the SCMD dataset) to volume in DDDs.

* SCMD volumes data is provided in vmp quantity - this means different things for different products. e.g. 100mg powder = 1 vmp quantity but 100mg/20ml solution for injection = 20 vmp quantity, even though both VMPS have same strength of ingredient.

* To translate volume in VMP quantity to volume in DDDs we need to go through a few steps:

  + First, translate volume in VMP quantity to volumes in singles of the product (i.e. number of vials)

  + Then translate volume in singles of product to volume in strength of ingredient (i.e. number of mgs of active ingredient)
  
  + Finally translate volume in strength of ingredient to volume in DDDs, using the DDD information published by the WHO. 

```{r, message = FALSE, warning = FALSE}
asthma_mabs_with_DDD <- asthma_mabs %>%
  mutate(volume_singles = total_quantity / udfs,
         volume_mg_strength = volume_singles * ifelse(is.na(strnt_dnmtr_val), strnt_nmrtr_val, strnt_nmrtr_val * (udfs / strnt_dnmtr_val)),
         mg_per_DDD = case_when(
           vtmnm == "Omalizumab" ~ 16,
           vtmnm == "Mepolizumab" ~ 3.6,
           vtmnm == "Reslizumab" ~ 7.1,
           vtmnm == "Benralizumab" ~ 0.54,
           TRUE ~ NA_real_
         ), 
         volume_DDD = volume_mg_strength / mg_per_DDD
  )

kable(asthma_mabs_with_DDD[1:5, ], caption = "Preview of asthma_mabs_with_DDD data table")
```

## Analysis

### Use of Asthma mAbs over time

The public secondary care medicines dataset provides information on hospital use of asthma mAbs from January 2019 to July 2020. 

At a national level we can see that the total use of asthma mAbs has increase over this period, but the size of the change varies by type of chemical. Use of Benralizumab has increased the most (in both absolute and proportional terms) and there is very little use of Reslizumab over the period.

```{r, message = FALSE, warning = FALSE}
Nat_time_series <- asthma_mabs_with_DDD %>%
  group_by(year_month, vtmnm) %>%
  summarise(volume_DDD = sum(volume_DDD))

ggplot(Nat_time_series, aes(x = year_month, y = volume_DDD, group = vtmnm)) +
  geom_line(aes(color = vtmnm), size = 2) +
  labs(x = "Month", y = "Total volume issued in hospitals - DDD") +
  scale_y_continuous(labels = scales::number) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90)) +
  ggtitle("Prescribing of asthma mAbs over time - National")
```

At a regional level it looks like the take up of asthma mAbs does vary by region - London and the North West seem to be the drivers of increase in use of Benralizumab with the North Easy and Yorkshire, the Midlands and the South East using more of Mepolizumab. There is a large variation in volume of use of Omalizumab between regions, with the North East and Yorkshire having the highest use over time and East of England having the lowest use.

```{r, message = FALSE, warning = FALSE, fig.width= 10, fig.height = 7}
Reg_time_series <- asthma_mabs_with_DDD %>%
  group_by(COMM_REGION_NAME, year_month, vtmnm) %>%
  summarise(volume_DDD = sum(volume_DDD))

ggplot(Reg_time_series, aes(x = year_month, y = volume_DDD, group = COMM_REGION_NAME)) +
  geom_line(aes(color = COMM_REGION_NAME), size = 1) +
  facet_wrap(facets=vars(vtmnm)) +
  labs(x = "Month", y = "Total volume issued in hospitals - DDD") +
  scale_y_continuous(labels = scales::number) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90),
        legend.position = "bottom") +
  ggtitle("Prescribing of asthma mAbs over time - Regional")
```

### Variation in the use of Asthma mAbs - by region and trust

We can look at the total issues of asthma mAbs over the latest 12 months of data available, August 2019 to July 2020 inclusive, to understand the variation in total issues and type of issues.

At a regional level the volume of total asthma mAbs issued in hospitals ranges from 141,000 DDDs in the East of England to 694,000 issues in the North East and Yorkshire. Across most regions Omalizumab and Mepolizumab make up the vast majority of issues, with the exception of London where Benralizumab has the greatest share of issues.

```{r, message = FALSE, warning = FALSE}
# Look at regional use by product
Reg_split <- asthma_mabs_with_DDD %>%
  filter(year_month >= as.Date("2019-08-01") & year_month <= as.Date("2020-07-01")) %>%
  group_by(COMM_REGION_NAME, vtmnm) %>%
  summarise(volume_DDD = sum(volume_DDD)) %>%
  group_by(COMM_REGION_NAME) %>%
  mutate(prop_use = volume_DDD / sum(volume_DDD),
         pos = cumsum(volume_DDD) - volume_DDD/2,
         total = sum(volume_DDD)) %>%
  ungroup() %>%
  arrange(desc(total))

ggplot(data = Reg_split, aes(x = reorder(COMM_REGION_NAME,total), y = volume_DDD, fill = vtmnm)) +
  geom_bar(position = "stack", stat = "identity") +
  theme_bw() +
  labs(x = "Region",
       y = "Volume of asthma mAbs issues - DDDs") +
  theme(strip.text.x = element_text(size = 8),
        panel.grid.major.x = element_blank(),
        legend.position = "bottom") + 
  scale_y_continuous(labels = scales::number) +
  ggtitle(label = "Regional use of asthma mAbs",
          subtitle = "Between August 2019 and July 2020")
```

At a trust level, again there is clear variation in the total issues of asthma mAbs between August 2019 and July 2020 and in the type of asthma mAb used. For trusts with very low issues Omalizumab seems to be the preferred choice, whereas there is more variation in the type of asthma mAb used in trusts with higher volumes of issues.

```{r, message = FALSE, warning = FALSE}
# Look at trust use by product
Trust_split <- asthma_mabs_with_DDD %>%
  filter(year_month >= as.Date("2019-08-01") & year_month <= as.Date("2020-07-01")) %>%
  group_by(trust_name, ods_code, vtmnm) %>%
  summarise(volume_DDD = sum(volume_DDD)) %>%
  group_by(trust_name, ods_code) %>%
  mutate(prop_use = volume_DDD / sum(volume_DDD),
         pos = cumsum(volume_DDD) - volume_DDD/2,
         total = sum(volume_DDD)) %>%
  ungroup() %>%
  arrange(desc(total))

# stacked bar chart
ggplot(data = Trust_split, aes(x = reorder(ods_code,total), y = volume_DDD, fill = vtmnm)) +
  geom_bar(position = "stack", stat = "identity") +
  theme_bw() +
  labs(x = "Trust",
       y = "Volume of asthma mAbs issues - DDDs") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),
        strip.text.x = element_text(size = 8),
        panel.grid.major.x = element_blank(),
        legend.position = "bottom") + 
  scale_y_continuous(labels = scales::number) +
  ggtitle(label = "Trust use of asthma mAbs",
          subtitle = "Between August 2019 and July 2020")

# scatter plot - total asthma mAb prescribing by % Omalizumab
trust_volume_by_prop_Omal <- asthma_mabs_with_DDD %>%
  filter(year_month >= as.Date("2019-08-01") & year_month <= as.Date("2020-07-01")) %>%
  mutate(vol_omalizumab = ifelse(vtmnm == "Omalizumab", volume_DDD, 0)) %>%
  group_by(ods_code) %>%
  summarise(volume_DDD = sum(volume_DDD),
            prop_omalizumab = sum(vol_omalizumab) / sum(volume_DDD)) %>%
  arrange(desc(volume_DDD))

ggplot(trust_volume_by_prop_Omal, aes(x = volume_DDD, y = prop_omalizumab)) +
  geom_point(size = 2, alpha=0.2, shape=16) +
  labs(x = "Total asthma mAb issues in DDDs", y = "% of total volume through Omalizumab") +
  scale_x_continuous(labels = scales::number) +
  scale_y_continuous(labels = scales::percent) +
  theme_bw() +
  ggtitle(label = "Comparing prescribing of Asthma mAbs and % of prescribing through Omalizumab",
          subtitle = "Trust level - Between August 2019 and July 2020")

```

### Top twenty trusts

Looking at the twenty trusts that use the greatest volume fo asthma mAbs we can see that University of Manchester and Guy's and St Thomas' trusts are the outliers in terms of total usage and Royal Brompton & Harefield trust stands out because there is no use of Omalizumab.

```{r, message = FALSE, warning = FALSE, fig.width = 10}
# Top 20 trusts
Trust_split_Top20 <- Trust_split %>%
  mutate(rank = dense_rank(desc(total))) %>%
  arrange(desc(total)) %>%
  filter(rank <= 20) %>%
  mutate(trust_name = as.character(trust_name),
         trust_name = case_when(ods_code == "RM2" ~ "UNIVERSITY HOSPITAL OF SOUTH MANCHESTER NHS FOUNDATION TRUST",
                                ods_code == "RR1" ~  "HEART OF ENGLAND NHS FOUNDATION TRUST",
                                TRUE ~ trust_name)
  )

ggplot(data = Trust_split_Top20, aes(x = reorder(trust_name,total), y = volume_DDD, fill = vtmnm)) +
  geom_bar(position = "stack", stat = "identity") +
  theme_bw() +
  labs(x = "Trust",
       y = "Volume of asthma mAbs issues - DDDs") +
  theme(panel.grid.major.y = element_blank(),
        legend.position = "bottom") + 
  scale_y_continuous(labels = scales::number) +
  ggtitle(label = "Trust use of asthma mAbs - Top 20 trusts by DDD volume",
          subtitle = "Between August 2019 and July 2020") +
  coord_flip()
```
